---
title: "EdEx Notes"
author: "Derek Hall"
date: "4/29/2020"
output: pdf_document
---

# Indexing Functions

## *which*

*which* gives the entries of a vector that are true. 
e.g. 

```{r}
x <- c(F, T, F, T, T, F)

which(x)
```

This function allows you to create a smaller index object since you only store the index of the variables of interest. 

## *match*

*match* outputs what indices of the second variable that match the entries of the first variable. 

e.g. 

```{r}
match(c("a", "b", "y", "z"), letters)
```

## *%in%*

*%in%* allows you to return a logical vector on whether the elements in a first variable are in a second variable. 

```{r}

c("a", "b", 1) %in% letters

```

*%in%* can be combined with *which* to generate an index of the first variable that are present in the second. 

```{r}
x <- c("a", "b", 1)

ind <- which(x %in% letters)

print(c(ind, x[ind]))

```

# The dplyr package

The dplyr package (part of tidyverse) can be used to intuitively manipulate data tables. 

main functions:


- *mutate*: Used to add a new column or change an existing one.   
- *filter*: Used to subset the data by rows.  
- *select*: Used to subset the data by columns.  
- *%>%* (Pipe Operator): Sends the results of one function to another function to perform multiple manipulations in sequence.  

```{r}
df <- data.frame(names = c("Suzy", "Becky", "Joan"), Apples = c(1, 4, 8), Shifts = c(1, 2, 4))

mutate(df, rate = Apples/Shifts)
```

