---
title: "R Programming Week 3 Notes"
author: "Derek Hall"
date: "3/25/2020"
output: html_document
---

# Looping Commands

- *lapply*: loop over a list and evaluate a f(x) on each element  
- *sapply*: lapply but with simplified results  
- *apply*: apply a function over the margins of an array  
- *tapply*: apply a function over subsets of a vector
- *mapply*: multivariate version of lapply  
- *split*: auxilary function useful in conjection with *lapply*.

## lapply

Always returns a list. If the input is not a list, it will be coerced into list form (with *as.list*)

*e.g.*

```{r} 
x <- list(a = 1:5, b = rnorm(10))

lapply(x, mean)

```

```{r}
x <- 1:4

lapply(x, runif)

lapply(x, runif, min = 0, max = 10)
```

*lapply* (and the other apply functions) make heavy use of anonymous functions. Anonymous functions are functions that don't have names and are not assigned in the environment. 

```{r}
x <- list(a = matrix(1:4, 2, 2), b = matrix(1:6, 3, 2))
x

lapply(x, function(elt) elt[, 1])
```

## sapply

Will try to simplify the result of *lapply* if possible. 

- If the result is a list where every element is length 1, then a vector is returned  
- If the result is a list where every element is a vector of the same length (>1), a matrix is returned
- If it can't figure it out, a list is returned 

## apply

Evaluates a function (often an anonymous one) over the margins of an array. 

*Note:* apply is not faster than writing a for loop, it is just less typing. 

*margins*: chooses which dimension to take for the function. *e.g.* for a matrix, rows are dim 1 and col. are dim 2. 

```{r}
x <- matrix(rnorm(200), 20, 10)

apply(x, 2, mean) #the mean of the columns

apply(x, 1, sum) #the sum of the rows
```

rowSums, rowMeans, colSums, and colMeans are equivalent to simple apply functions, but are much faster if the matrix is large. 

apply can also be used on multidimensional matrices 

```{r}
a <- array(rnorm(2 * 2 * 10), c(2, 2, 10))
apply(a, c(1, 2), mean)

rowMeans(a, dims = 2)
```

## mapply

A multivariate version of apply. Applies a function in parallel over a set of arguments. 

```{r}
str(mapply)

```

- FUN is a function to apply  
- ... contains arguments to apply over.  
- MoreArgs is a list of other arguments to FUN  
- SIMPLIFY indicates whether the result should be simplified. 

*e.g.* The following is tedious to type 

list(rep(1, 4), rep(2, 3), rep(3, 2), rep(4, 1))

instead we can do

```{r}
mapply(rep, 1:4, 4:1)
```

## tapply 

Used to apply a function over subsets of a vector. 

```{r}
str(tapply)
```

- X is a vector  
- INDEX is a factor or a list of factors (if not factors, will be coerced to factors)
- FUN is a function to be applied  
- ... contains other arguments to be passed to FUN
- simplify, same as above

*e.g.* Take group means.

```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10)
f

tapply(x, f, mean)
```

In the above example, x is a list of three sets of 10 random numbers are generated. f is an index variable that indicates the first 10 variables belong to group 1, the second 10 belong to group 2, and the third 10 belong to group three. By passing x and f to tapply, the means are calculated using the f index to assign to groups 1, 2, and 3. 

Of particular interest, when working with a data frame, tapply can be passed subsetted column (or row) vectors as the list to be passed to the function and the index list. 

*e.g*

```{r}
df <- read.csv("D:/Users/derek/OneDrive - The Ottawa Hospital/Projects/01 WntPCP Ciliogenesis Crosstalk EVF/01_DataAnalysis/data/P01_rawdata_ciliatimecourse.csv")

tapply(df$pax7pos_cilianeg, df$time_hours, mean)

```

## split

Takes a vector or other objects and splits it into groups determined by a factor or list of factors

```{r}
str(split)
```

x = vector or list or data frame
f = factor or a list of factors
drop = indicates whether empty factor levels should be dropped. 

```{r}
x <- c(rnorm(10), runif(10), rnorm(10,1))
f <- gl(3, 10)
split(x, f)

```

*Note*: the *gl* function is super useful for generating a factor list level for subsetting another variable. 

```{r}
str(gl)
```

n = number of levels
k = number of replicates per level
length = how long the list repeats for
labels = optional list of labels to apply to the levels. 
order = define the order of the levels 

split is useful for dividing complicated data frames. 

```{r}
library(datasets)
head(airquality)
```

```{r}
s <- split(airquality, airquality$Month)
lapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")])}
       )
```

```{r}
s <- split(airquality, airquality$Month)
sapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)}
       )
```

**Splitting more than one level**: Sometimes, you might want to split on more than two levels (e.g. gender and treatment). To do so, use the *interaction* function


```{r}
x <- rnorm(10)
f1 <- gl(2, 5)
f2 <- gl(5, 2)
interaction(f1, f2)

```

```{r}
str(split(x, list(f1, f2)))
```

```{r}
str(split(x, list(f1, f2), drop = TRUE))
```

# Debugging

Several types of "error messages"

- message: a generic notification. The function still continues.  
- warning: an indication something is wrong but not fatal. i.e. something unexpected happened.  
- error: An indication that a fatal problem occured; execution stops, produced by the *stop* function
- condition: a generic concept for indicating that something unexpected occured. All of the above are conditions.  

warning: 

```{r}
log(-1)
```


```{r}
printmessage <- function(x) {
        if(x > 0)
                print("x is greater than zero")
        else
                print("x is less than or equal to zero")
        invisible(x)
}

#printmessage(NA)
```

*note*: invisible allows you to hide the results of a function if you don't want them printing to the console. 

```{r}
printmessage2 <- function(x) {
        if(is.na(x))
                print("x is a missing value!")
        else if(x > 0)
                print("x is greater than zero")
        else
                print("x is less than or equal to zero")
        invisible(x)
}

printmessage2(NA)
```

Answer these questions if something is wrong:

- What was your input? How did you call the function?  
- What were you expecting? Output messages, other results?
- What did you get?  
- How does what you got differ from what you were expecting?  
- Were your expectations correct in the first place?  
- Can you reproduce the problem (exactly)?  

## Debugging tools in R

- *traceback*: prints the function call stack after an error occurs. Tells you how far in you were when the error occurred. 
- *debug*: flags a function for "debug" mode, which allows you to step through the execution of a function one line at a time. 
- *browser*: suspends the execution of a function wherever it is called and puts the function in debug mode. Useful if you don't need to debug the whole function, just a section of it.  
- *trace*: allows you to insert debugging code into a function at specific places. 
- *recover*: allows you to modify the error behavior so that you can browse the function call stack. 

### Traceback 

Useful for sharring issues with other users. Gives the list of function calls. Must be called immediately after an error. 

```{r}
mean(x)
```

```{r}
traceback(
)
```

lm(y - x)

```{r}
traceback()
```

### Debug and Browse

*Debug* Creates a separate *browse* environment, which is like a workspace within a workspace. type "n" to move to the next line of code. Continues until you hit the error. 

If you have called debug on multiple functions that are in a call stack, when the Browser reaches the second (third, fourth, etc...) function it will open a browser within a browser (nesting). The current level is indicated in [] brackets next to "Browse" in the command line. 

### Recover

"options(error = recover)" will set R to recover mode until you close the workspace. 

When you hit an error, you will automatically get a traceback of the error. 

Gives you an ordered list of the function call stack from the traceback, that you can navigate through with the console. Useful for large nested functions. 
