---
title: "quiz_week2"
author: "Derek Hall"
date: "3/20/2020"
output: pdf_document
---

# Question 1 
Suppose I define the following function in R

```{r}
cube <- function(x, n) {
        x^3 
}
```

What is the result of running
```{r}
cube(3)
```

Functions are assessed lazily so the n will be ignored until it is needed. 

# Question 2

The following code will produce a warning in R. Why?

```{r}
x <- 1:10
if(x > 5) {
        x <- 0
}
```
If can only test a single logical value. If the goal is to set all values in the list x that are above 5 to 0, the If statement must be looped for each value in the x vector. This can be acheived with a for loop. 

```{r}
x <- 1:10
for(i in x) {
        if(i > 5) {
                x[i] <- 0
        }
}

x
```


# Question 3

Consider the following function
```{r}
f <- function(x) {
        g <- function(y) {
                y + z
        }
        z <- 4
        x + g(x)
}
```

If I then run in R what value is returned?
```{r}
z <- 10
f(3)
```

g is defined within function f. Therefore, z will be called from the f fuction environment and set at 4. (x + g(x)) = (3 + g(3)) = (3 + (3+4)) = 10

# Question 4

Consider the following expression. What is the value of 'y' after evaluating this expression? 

```{r}
x <- 5
y <- if(x < 3) {
        NA
} else {
        10
}
```

5 > 3, therefore y will be 10. 

```{r}
y
```

# Question 5

Consider the following R Function

```{r}
h <- function(x, y = NULL, d = 3L) {
        z <- cbind(x, d)
        if(!is.null(y))
                z <- z + y
        else
                z <- z + f
        g <- x + y / z
        if(d == 3L)
                return(g)
        g <- g + 10
        g
}
```

f is not defined by any function arguments and so is a free variable. 
