---
title: "quiz_week3"
author: "Derek Hall"
date: "4/1/2020"
output: pdf_document
---

# Question 1

In this dataset, what is the mean of 'Sepal.Length' for the species virginica? Please round your answer to the nearest whole number.

```{r}
library(datasets)
data(iris)
?iris


tapply(iris$Sepal.Length, iris$Species, mean)


```

# Question 2

Continuing with the 'iris' dataset from the previous Question, what R code returns a vector of the means of the variables 'Sepal.Length', 'Sepal.Width', 'Petal.Length', and 'Petal.Width'?

```{r}
apply(iris[,1:4], 2, mean)
```

# Question 3

How can one calculate the average miles per gallon (mpg) by number of cylinders in the car (cyl)? Select all that apply.(in the mtcars dataset)

```{r}
data(mtcars)
head(mtcars)

tapply(mtcars$mpg, mtcars$cyl, mean)
with(mtcars, tapply(mpg, cyl, mean))
sapply(split(mtcars$mpg, mtcars$cyl), mean)

```

# Question 4

Continuing with the 'mtcars' dataset from the previous Question, what is the absolute difference between the average horsepower of 4-cylinder cars and the average horsepower of 8-cylinder cars?

```{r}
x <- tapply(mtcars$hp, mtcars$cyl, mean)
abs(x[1] - x[3])
```

